---
title: "ANLY560: Lab 1"
author: "Dr. Purna Gamage"

output: rmdformats::robobook
---

```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(lubridate)
```

# Time Series Object

## How to create a Time Series in R ?

1. You can create time series data using `ts()` function. The function `ts()` is used to create time-series objects.

The format is `ts(vector, start=, end=, frequency=)`

The function takes three arguments:

  - a numeric vector of data
  - an optional start date, specified using the start parameter
  - an optional frequency, specified using the frequency parameter
  
 the start date (in the format c(year, month, day) or c(year, quarter)), and the frequency (e.g., 12 for monthly data or 4 for quarterly data..etc). 
 
 2. Defining Frequency;

The value of argument frequency is used when the series is sampled an integral number of times in each unit time interval. 

  - For example, one could use a value of 7 for frequency when the data are sampled daily, and the natural time period is a week, 

  - or 12 when the data are sampled monthly and the natural time period is a year. 

  - Values of 4 and 12 are assumed in (e.g.) print methods to imply a _quarterly_ and monthly _series_ respectively.
  
  More detais: <https://robjhyndman.com/hyndsight/seasonal-periods/>
  
### Example 2.1

Here is an example of creating a monthly time series:
  
```{r}
data <- c(1, 2, 3, 4, 5)
start_date <- c(2023, 1)
frequency <- 12
time_series <- ts(data, start = start_date, frequency = frequency)
time_series
```

 
 In this example, the data is a vector of numbers, the start date is January 2010, and the frequency is 12 (monthly). The `ts()` function creates a time series object, which can be plotted or used for other analyses.


   - `as.ts()` and `is.ts()` coerce an object to a time-series and test whether an object is a time series.

### Example 2.2

```{r}
purely_random_process =as.ts(rnorm(1000))

autoplot(purely_random_process) +ggtitle("A pure random Process - White Noise")
is.ts(purely_random_process) #TRUE
```

### Example 2.3: 

In this case the age of death of 42 successive kings of England has been read into the variable ‘kings’; starting with William the Conqueror.

```{r}
kings <- scan("http://robjhyndman.com/tsdldata/misc/kings.dat",skip=3)
kingstimeseries <- ts(kings)
kingstimeseries

autoplot(kingstimeseries)+ggtitle("Age of death of 42 successive kings of England")
```

```{r}

is.ts(kings)
is.ts(kingstimeseries)

```


   - Alternatively, you can use the `zoo` or `xts` packages that provides more flexibility in handling time series data.

### Example 2.4

```{r}
library(zoo)
library(xts)
data <- c(1, 2, 3, 4, 5)
dates <- as.Date(c("2010-01-01", "2010-02-01", "2010-03-01", "2010-04-01", "2010-05-01"))
time_series <- xts(data, order.by = dates)
time_series
str(time_series)
```


Here, the data is a vector of numbers and dates is a vector of dates. The `xts()` function creates an `xts` object with dates as the index and data as the values.
  
   
   
3. You can also specify the first year that the data was collected, and the first interval in that year by using the `‘start’` parameter in the `ts()` function. For example, if the first data point corresponds to the second quarter of 1986, you would set `start=c(1986,2)`.

(For imported data: Upon importing your data into R, use `ts()` function as follows. The input Data used here is ideally a numeric vector of the class `‘numeric’` or `‘integer’`.)   

### Example 3.1

```{r}

# save a numeric vector containing 72 monthly observations
# from Jan 2009 to Dec 2014 as a time series object
myts <- ts(1:72, start=c(2009, 1), end=c(2014, 12), frequency=12)
myts

# subset the time series (June 2014 to December 2014)
myts2 <- window(myts, start=c(2014, 6), end=c(2014, 12))
myts2


# plot series
#plot(myts)

```
### Example 3.2

```{r}

ts(1:10, frequency = 4, start = c(1959, 2)) # start from 2nd Quarter of 1959
# frequency 4 => Quarterly Data

ts (1:10, frequency = 12, start = 1990) # freq 12 => Monthly data. Start from Jan 1990 to Oct 10. No need to specify the end 

z <- ts(matrix(rnorm(300), 100, 3), start = c(1961, 1), frequency = 12) #generating 3 random series
class(z)
head(z) # as "matrix"

######## When you import data #######################
#ts (inputData, frequency = 4, start = c(1959, 2)) # frequency 4 => Quarterly Data
#ts (inputData, start=c(2009), end=c(2014), frequency=1) # Yearly Data

#####################################################
```


### Example 3.3: 

the number of births per month in New York city, from January 1946 to December 1959.

<https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html>
   
   
```{r}
births <- scan("http://robjhyndman.com/tsdldata/data/nybirths.dat")
head(births)
is.ts(births)


### So you need to convert this to ts object

#### converting to a Time series object: ######
birthstimeseries <- ts(births, frequency=12, start=c(1946,1))

is.ts(birthstimeseries)

autoplot(birthstimeseries) +ggtitle("Births per month in New York city, from January 1946 to December 1959")
```

### Example 3.4: 

monthly sales for a souvenir shop at a beach resort town in Queensland, Australia, for January 1987-December 1993. 

```{r}
souvenir <- scan("http://robjhyndman.com/tsdldata/data/fancy.dat")

#### converting to a Time series object: ######
souvenirtimeseries <- ts(souvenir, frequency=12, start=c(1987,1))
souvenirtimeseries
```


### Example 3.5

Percent Change From Preceding Period in Real Gross Domestic Product: GDP in USA from 1980 to 2022.

```{r}
gdp=read.csv("gdp_2022.csv") #1980 Q1 to 2022 Q3
head(gdp)
colnames(gdp)<-c("GDP")

#### converting to a Time series object: ######
gdp.ts=ts(gdp$GDP, start=c(1980,1), end=c(2022,3), frequency = 4)

autoplot(gdp.ts)+ggtitle("GDP in USA from 1980 to 2022")
```

** Let's plot in `Plotly` to identify the time of exact drop in the percentage **

  - plot in plotly
  - I need a dataframe format to use plotly. Therfore, I need a date column.
  
1. This is one way of creating a date column. But I prefer method 2.

```{r}

### This is one way
#library(zoo)

# Create a sequence of quarters from 1980Q1 to 2022Q3
#quarters <- seq(as.yearqtr("1980Q1"), as.yearqtr("2022Q3"), by = "quarter")

# Create a time series object with the quarters as the time index and random values as the data
#data <- zoo(rnorm(length(quarters)), quarters)
```

Here, the `seq()` function creates a sequence of quarters starting from `1980Q1` and ending at `2022Q4` with a step of one quarter. The `as.yearqtr()` function is used to convert the string representation of the quarters to the format understood by the `seq()` function.

The `zoo()` function creates a zoo object with the quarters as the time index and random values generated by `rnorm()` as the data.

You can use this time series object for further analysis or visualization.

2. Creating Date column using `xtx` package.

Alternatively, you can use the `xts` package to create a time series dataset with quarters as the index.

```{r}

#library(xts)

# Create a sequence of quarters from 1980Q1 to 2022Q4
quarters <- seq(as.Date("1980-01-01"), as.Date("2022-07-01"), by = "quarter")

# Create an xts object with the quarters as the time index and random values as the data
#this is a simulation example;
#data <- xts(rnorm(length(quarters)), order.by = quarters)


#length(gdp$GDP) 
length(quarters) 
start(quarters)
end(quarters)

gdp_data=data.frame(gdp$GDP,quarters)

```

Here, the `seq()` function creates a sequence of quarters starting from `1980-01-01` and ending at `2022-07-01` with a step of three months, which corresponds to one quarter. The `xts()` function creates an `xts` object with the quarters as the time index and random values generated by `rnorm()` as the data.

```{r}
colnames(gdp_data) = c("GDP","date")
head(gdp_data)
str(gdp_data) # date is in date format


fig <- plot_ly(gdp_data, x = ~date, y = ~GDP, type = 'scatter', mode = 'lines')

fig <- fig %>% layout(title = "GDP in US from 1980 to 2022")
fig
```


3. Using `tsibble` package to create a date column.

You can also use the `tsibble` package, which provides a more modern way to work with time series data by providing a dataframe-like structure.

The tsibble package provides a data infrastructure for tidy temporal data with wrangling tools. Adapting the tidy data principles, tsibble is a data- and model-oriented object. 

More Details : <https://tsibble.tidyverts.org/#:~:text=The%20tsibble%20package%20provides%20a,ordering%20from%20past%20to%20present.>

```{r, message=FALSE, warning=FALSE}
library(tsibble)
data <- tibble(
  date = seq(as.Date("1980-01-01"), as.Date("2022-07-01"), by = "quarter"),
  value = gdp_data$GDP
)  %>%
  as_tsibble(index = value, key = date)

colnames(data) = c("date","GDP")
#library(plotly)
fig <- plot_ly(data, x = ~date, y = ~GDP, type = 'scatter', mode = 'lines')

fig <- fig %>% layout(title = "GDP in US from 1980 to 2022")
fig
```

Here, the `seq()` function creates a sequence of quarters starting from `1980-01-01` and ending at `2022-12-31` with a step of three months, which corresponds to one quarter. `tibble()` creates a data frame with two columns, 'date' and 'value'. The `as_tsibble()` function creates a `tsibble` object with the date as the time index, and GDP values as the data.


### Example 3.6

Actual GDP is usually increasing over the years. 

```{r}
GDP<-read.csv("GDP.csv",header = TRUE)
head(GDP)

#make a time series object
GDP_ts <- ts(GDP$GDP,star=decimal_date(as.Date("1947-01-01")), frequency = 4)
p<-autoplot(GDP_ts,color = 'blue')+ggtitle("GDP in US from 1947 to 2021")
ggplotly(p)
```


Gross domestic product (GDP) is the value of all goods and services produced within a country over a certain period of time. GDP generally increases over time due to a combination of population growth, inflation, and economic growth. Economic growth occurs when there is an increase in the production of goods and services, which can be driven by factors such as technological advancements, increased productivity, and foreign investment. Additionally, GDP can also increase as a result of government policies aimed at stimulating economic activity, such as increased spending on infrastructure or tax cuts.

# Date Time Format

### Example 4.1

 In this example, first load the `lubridate` and `zoo` libraries to work with dates and times. 

 Then create a sample time series dataset with a date column in `"YYYY-MM-DD"` format and a value column. 
 
 Here `as.POSIXct()` function is used to convert the date column to a `POSIXct` format and `zoo()` function to create a time series object. 
 
 Next, change the format of the date/time column using the `format()` function, and then extracted the year, month, and day using the `year()`, `month()`, and `day()` functions from the `lubridate` library.

 You can try different date/time formats and use the appropriate functions to extract different parts of the date/time, depending on your analysis needs.

```{r, eval=FALSE}
# Load the required libraries
library(lubridate)
library(zoo)

# Create a sample time series data set
date_range <- as.Date(c("2020-01-01", "2020-01-02", "2020-01-03", "2020-01-04"))
values <- c(1, 2, 3, 4)
ts_data <- data.frame(date_range, values)
ts_data 

# Convert the date column to a POSIXct format
ts_data$date_range <- as.POSIXct(ts_data$date_range)

# Create a time series object using the 'zoo' library
ts_object <- zoo(ts_data$values, order.by = ts_data$date_range)

# Print the time series object
print(ts_object)

#> "POSIXct" "numeric"
#> 2020-01-01 1
#> 2020-01-02 2
#> 2020-01-03 3
#> 2020-01-04 4

# Change the format of the date/time column
index(ts_object) <- format(index(ts_object), format = "%Y-%m-%d %H:%M:%S")

# Print the time series object
print(ts_object)
#or
#ts_object

#> "2020-01-01 00:00:00" 1
#> "2020-01-02 00:00:00" 2
#> "2020-01-03 00:00:00" 3
#> "2020-01-04 00:00:00" 4

# Extract the year, month, and day from the date/time column
year <- lubridate::year(index(ts_object))
month <- month(index(ts_object), label = TRUE)
day <- day(index(ts_object))

```




## as.Date()

The `as.Date()` function in R is used to convert a character or numeric value into a date object. It is a part of the base R package and does not require any additional packages to be loaded.

### Example 4.2:

Here is an example of how you can use the `as.Date()` function:

```{r}
# Convert a character string to a date object
date_string <- "2022-01-01"
date_object <- as.Date(date_string, format = "%Y-%m-%d")

# The date object is of class "Date"
class(date_object)

# Output: "Date"

# Print the date object
print(date_object)

# Output: "2022-01-01"

```

In this example, the `as.Date()` function converts the character string `"2022-01-01"` into a date object. The format parameter is used to specify the format of the input string. 

### Example 4.3:

You can also convert numeric values to date objects using the `as.Date()` function.

```{r}
# Convert a numeric value to a date object
numeric_value <- 17401
date_object <- as.Date(numeric_value, origin = "1970-01-01")

# The date object is of class "Date"
class(date_object)

# Output: "Date"

# Print the date object
print(date_object)

# Output: "1970-01-18"

```

Here, the `as.Date()` function converts the numeric value 17401 into a date object with an origin of `1970-01-01`. The origin parameter is used to specify the reference date for numeric values.

In the above example, the numeric value 17401 represents the number of days since 1970-01-01, so the date object is equivalent to 1970-01-18.

Note that the `as.Date()` function returns a date object that only stores the date and not the time. If you want to work with date and time, you can use the `as.POSIXct()` or `as.POSIXlt()` functions to convert character or numeric values to `POSIXct` or `POSIXlt` objects respectively.

Here is another example:

### Example 4.4: Apple Stock Prices

Forecasting stock prices is a very difficult and challenging task in the financial market because the trends of stock prices are non-linear and non-stationary time-series data.

We have been tasked with forecasting the price of the AAPL(Apple) stock index and have
been provided with historical data. This includes features like opening and closing stock prices, volume, date, and so on. As a data scientist, which kind of problem would you classify this as? Time series modeling, of course.

_**Hypotheses that affect the closing price(dependent variable for this time series problem):**_

1. There will be an increase in the closing price from September to October 2019.

_Explanation — iphone11 came out on September 20, 2019, so we would expect the stock
price to increase during the next two months._

2. The closing prices increase from November to December every year.

_Explanation — November and December months are generally considered as holiday
season due to thanksgiving and Christmas. Customers generally spend more during
holiday season._

_"Investors should purchase stocks like they purchase groceries, not like they purchase perfume." -Ben Graham_

```{r}
data=read.csv("AAPL3.csv")
head(data)
close=data$Close

## convert to time series object
xt=ts(close, star=decimal_date(as.Date("2014-03-03")), frequency = 365.25)

```



# More Examples

## Example: 5.1 More Stocks

```{r}
options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

tickers = c("AAPL","TSLA","AMZN" )
for (i in tickers){
  getSymbols(i,
             from = "2012-10-01",
             to = "2022-12-01")}

x <- list(
  title = "date"
)
y <- list(
  title = "value"
)

stock <- data.frame(AAPL$AAPL.Adjusted,
                    TSLA$TSLA.Adjusted,
                    AMZN$AMZN.Adjusted)


stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append(tickers,'Dates')

stock$date<-as.Date(stock$Dates,"%Y-%m-%d")
head(stock)


```

### Apple Stocks

```{r}
pa<-ggplot(stock, aes(x=date)) +
  geom_line(aes(y=AAPL, colour="AAPL"))+
ggtitle("Adjusted Closing Price for Apple Inc. from 2012 to 2021")

ggplotly(pa)
```


### AMZN Stocks

```{r}
AMZN[1,]
AMZN[nrow(AMZN),]

AMZN.close=AMZN$AMZN.Adjusted
AMZN.close=ts(AMZN.close, star=decimal_date(as.Date("2012-10-01")),frequency = 365.25)
str(AMZN.close)

pz<-autoplot(AMZN.close, colour = 'blue')+ggtitle("AMZN Stock market closing prices from 2012 to 2021")

ggplotly(pz)
```

### TSLA Stocks

```{r}

fig <- plot_ly()%>%
  add_trace(data = stock, type = 'scatter', mode = 'lines', fill = 'tozeroy', x = ~date, y = ~TSLA, name = 'TSLA')%>%
  layout(showlegend = F, yaxis = list(range = c(10,450),
                                      gridcolor = 'ffff'),
         xaxis = list(gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6')
options(warn = -1)
fig <- fig %>%
  layout(
    xaxis = list(gridcolor = 'ffff'),
    yaxis = list(gridcolor = 'ffff'),
    plot_bgcolor='#e5ecf6', width = 500)


fig
```


### Apple Vs. AMZN Vs TSLA

```{r}
g4<- ggplot(stock, aes(x=date)) +
  geom_line(aes(y=AAPL, colour="AAPL"))+
  geom_line(aes(y=AMZN, colour="AMZN"))+
  geom_line(aes(y=TSLA, colour="TSLA"))+
   labs(
    title = "Stock Prices for the Tech Companies",
    subtitle = "From 2012-2022",
    x = "Date",
    y = "Adjusted Closing Prices")+
    guides(colour=guide_legend(title="Tech Companies")) 


ggplotly(g4) %>%
  layout(hovermode = "x")
```




## Example 5.2: Weather Forecast

```{r}
wt = read.csv("climate.csv", header = TRUE)
head(wt)

tobb= wt[1095:1344,c(3,11)]
head(tobb)
tobb$DATE<-as.Date(tobb$DATE,"%Y-%m-%d")

pw<-ggplot(tobb, aes(x=DATE, y=TOBS)) +
  geom_line(colour="blue") + 
  ggtitle("Temperature at the time of observation from 2021-01-01 to 2021-09-30 in Washington DC")

ggplotly(pw)
```

<https://www.ncdc.noaa.gov/cdo-web/>




# Interpreting time series plots

### Example 6.1: weekly economy passenger load on Ansett Airlines between Australia’s two largest cities.

```{r}
p<-autoplot(melsyd[,"Economy.Class"]) +
  ggtitle("Economy class passengers: Melbourne-Sydney") +
  xlab("Year") +
  ylab("Thousands")

ggplotly(p)
```

The time plot immediately reveals some interesting features.

  - There was a period in 1989 when no passengers were carried — this was due to an industrial dispute.

  - There was a period of reduced load in 1992. This was due to a trial in which some economy class seats were replaced by business class seats.

  - A large increase in passenger load occurred in the second half of 1991.
There are some large dips in load around the start of each year. These are due to holiday effects.

  - There is a long-term fluctuation in the level of the series which increases during 1987, decreases in 1989, and increases again through 1990 and 1991.

  - There are some periods of missing observations.

  - Any model will need to take all these features into account in order to effectively forecast the passenger load into the future.


# Gaussian White Noise

A particularly useful white noise series is Gaussian white noise, wherein the $w_t$ are independent normal random variables, with mean 0 and variance $\sigma ^2_w$ ; or more succinctly, 
$$wt \sim iid N(0, \sigma ^2_w)$$

White noise is an important concept in time series analysis and forecasting.

It is important for two main reasons:

1. Predictability: If your time series is white noise, then, by definition, it is random. You cannot reasonably model it and make predictions.

2. Model Diagnostics: The series of errors from a time series forecast model should ideally be white noise. Why?  
When forecast errors are white noise, it means that all of the signal information in the time series has been harnessed by the model in order to make predictions. All that is left is the random fluctuations that cannot be modeled.  

A sign that model predictions are not white noise is an indication that further improvements to the forecast model may be possible.


```{r}
wn2= rnorm(1000,0,1) #variance 1
autoplot(ts(wn2), ylab="White Noise")+ggtitle("Gaussian White Noise")
```

# Decomposition

## DECOMPOSE( ) and STL(): Time Series Decomposition in R

To make life easier, some R packages provides decomposition with a single line of code. As expected, our step-by-step decomposition (See Lab 1_Assignment for the full details and a demonstration on these steps) provides the same results as the DECOMPOSE( ) and STL( ) functions (see the graphs).

### Step 1: Import the Data 

* Additive 

```{r, message=FALSE, warning=FALSE}
# Additive 
# install.packages("fpp")
library(fpp)
data(ausbeer) # This ausbeer dataset is an example of additive time series 


timeserie_beer=ausbeer 

#pe<-autoplot(as.ts(timeserie_beer), ylab = "Beer production")+ggtitle("Full Quarterly Australian Beer production")
#ggplotly(pe)

# I only need the first part of the data for this analysis
timeserie_beer = tail(head(ausbeer, 17*4+2),17*4-4)

###  head(ausbeer, 17*4+2) shows me the data from 1956 Q1 to 1973 Q2
### From 1956 to 1973 : 17 years
### tail(head(ausbeer, 17*4+2),17*4-4) shows me the data from 1957 Q3 to 1973 Q2


autoplot(as.ts(timeserie_beer), ylab = "Beer production")

# As the metric values increase, the seasonality stays relatively constant.
```
```{r}
timeserie_beer=ausbeer 
ausbeer
#timeserie_beer

```

* Multiplicative 

```{r}
# Multiplicative 
# install.packages("Ecdat")
#install.packages("Ecfun")
#library(Ecdat)
#library(Ecfun)

data(AirPassengers) # This AirPassengers dataset is an example of multiplicative time series 
timeserie_air = AirPassengers

autoplot(as.ts(timeserie_air), xlab = 'Year', ylab = 'Passengers')+ggtitle('AirPassengers')
# The more passengers there are, the more seasonality is observed.
```

### Decompose() function 

* Additive 

The only requirement: seasonality is quarterly (frequency = 4)

```{r}
ts_beer = ts(timeserie_beer, frequency = 4)
decompose_beer = decompose(ts_beer, "additive")
autoplot(decompose_beer)

## Extract the ts components ##
autoplot(as.ts(decompose_beer$seasonal))
autoplot(as.ts(decompose_beer$trend))
autoplot(as.ts(decompose_beer$random))

```

* Multiplicative 


The classic Box & Jenkins airline data. Monthly totals of international airline passengers, 1949 to 1960.

You can see that the frequency is 12 here;

```{r}
frequency(AirPassengers)
```


```{r}
ts_air = ts(timeserie_air, frequency = 12)
decompose_air = decompose(ts_air, "multiplicative")
autoplot(decompose_air)

## Extract the ts components ##
autoplot(as.ts(decompose_air$seasonal))
autoplot(as.ts(decompose_air$trend))
autoplot(as.ts(decompose_air$random))

```


  - As we have seen before, the magnitude of the seasonal pattern is increasing. Therefore, we can use `multiplicative` decomposition.
  
  

### Example 7.1: lynx data

Annual number of lynx trapped in McKenzie river district of northwest Canada: 1821–1934.


```{r}
autoplot(lynx)+ggtitle("Lynx Population")
```

At first glance, the strong cycles in series might appear to make it non-stationary. But these cycles are aperiodic — they are caused when the lynx population becomes too large for the available feed, so that they stop breeding and the population falls to low numbers, then the regeneration of their food sources allows the population to grow again, and so on. In the long-term, the timing of these cycles is not predictable. Hence the series is stationary.


## STL( ) function:

```{r}
ts_beer = ts(timeserie_beer, frequency = 4)
stl_beer = stl(ts_beer, "periodic")

## Extract the ts components ##
seasonal_stl_beer   <- stl_beer$time.series[,1]
trend_stl_beer     <- stl_beer$time.series[,2]
random_stl_beer  <- stl_beer$time.series[,3]
 
autoplot(ts_beer) # original plot
autoplot(stl_beer) # decomposition

## Individual ts components ##
autoplot(as.ts(seasonal_stl_beer))
autoplot(trend_stl_beer)
autoplot(random_stl_beer)

```

### Example 8.1: Example for stl()

STL applied to the electrical equipment orders data.
With a larger window;

```{r}
autoplot(elecequip)+autolayer(ma(elecequip,12), series="12-MA")

elecequip %>%
  stl(t.window=13, s.window="periodic", robust=TRUE) %>%
  autoplot()+ggtitle("Electrical equipment manufacturing")
```


Here, it is more obvious that there has been a down-turn at the end of the series, and that the orders in 2009 were unusually low (corresponding to some large negative values in the remainder component).

The user must specify _s.window_ as there is no default. Setting it to be infinite is equivalent to forcing the seasonal component to be periodic (i.e., identical across years). 


  - Here I used 12-Moving average(MA) smoothing to smooth the series to see underline pattern.

  - Also, you can see after decomposition, the trend component looks much smoother. It is also can be seen that it is a combination of the cyclic and trend component.

  - Moreover, more randomness can be seen in the random component.
  



**With a smaller window;**

```{r}
elecequip %>%
  stl(t.window=5, s.window="periodic", robust=TRUE) %>%
  autoplot()+ggtitle("Electrical equipment manufacturing")
```

 - Here, after decomposition, the trend component looks less smoother with a smaller window. 

  - Moreover, more randomness can be seen in the trend component and less randomness in the random component.



## Why Decomposition?

** Decomposition is often used to remove the seasonal effect from a time series. 

** It provides a cleaner way to understand trends. 
For instance, lower ice cream sales during winter don’t necessarily mean a company is performing poorly. 

** To know whether or not this is the case, we need to remove the seasonality from the time series. 

** We also use the random (also call remainder) time series from the decomposed time series to detect anomalies and outliers.
